Lesson 3
========================================================

***


### What to Do First?
Notes:
Exploratory data analysis.
Plotting, summary data, etc.

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
setwd('/home/jay/classes/R.class')
pf = read.csv('pseudo_facebook.tsv', sep='\t')
names(pf)
summary(pf)
head(pf)

```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
qplot(data=pf, x=dob_day) +
  scale_x_discrete(breaks=1:31)
```

***

#### What are some things that you notice about this histogram?
Response:
The peak on day 1 of the month.
***

### Moira's Investigation
Notes:
Perception of audience size versus reality.
***

### Estimating Your Audience Size
Notes:


***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:
I posted something about a particular song recently.
#### How many of your friends do you think saw that post?
Response:
10
#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:
10%
***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:
facet_wrap and facet_grid
```{r Faceting}
qplot(data=pf, x=dob_day) +
  scale_x_discrete(breaks=1:31) + 
  facet_wrap(~dob_month, ncol=3)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:
Jan 1 has all those users that created the spike we saw earlier.
***

### Be Skeptical - Outliers and Anomalies
Notes:
What are the causes of the outliers?
Is it bad data, really just out there (extreme value)? 
How did it become bad if it is?
***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:
Her outlier was bad data about an extreme case.
***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(data=pf, x=friend_count)

```

#### How is this plot similar to Moira's first plot?
Response:
It has people at the higher end that causes the plot to obscure most of the data in the smaller counts.
***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
ggplot(data=pf, aes(x=friend_count)) +
  scale_x_continuous(limits=c(0, 1000), breaks=seq(0, 1000, 50))
```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:
```
qplot(x = friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```
### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 25) +
  facet_wrap(~gender) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

***

### Omitting NA Values
Notes: 
watch out for na.omit as that will omit any observations with na anywhere in them

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 25) +
  facet_wrap(~gender) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response: women

#### What's the difference between the median friend count for women and men?
Response: 22

#### Why would the median be a better measure than the mean?
Response: less impact due to big values

***

### Tenure
Notes:

```{r Tenure}
qplot(data=pf, x=tenure,
  color=I('black'), 
  fill=I('#099DD9'),
  binwidth=30)
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(data=pf, x=tenure/365,
  color=I('black'), 
  fill=I('#F79420'),
  binwidth=0.25) +
  scale_x_continuous(breaks=seq(1,7,1), lim=c(0,7))
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(data=pf, x=tenure/365,
  xlab="Years",
  ylab="Num Users",
  color=I('black'), 
  fill=I('#F79420'),
  binwidth=0.25) +
  scale_x_continuous(breaks=seq(1,7,1), lim=c(0,7))
```

***

### User Ages
Notes:

```{r User Ages}
summary(pf$age)
qplot(data=pf, x=age,
  xlab="Age",
  ylab="Num Users",
  color=I('black'), 
  fill=I('#F79420'),
  binwidth=5) + 
  #scale_x_discrete(breaks=seq(0,90,5), limit=c(0,90))
  scale_x_continuous(breaks=seq(0,90,5), limit=c(0,90))
```

#### What do you notice?
Response:
Some outliers in the upper range, probably people putting in fake birthdays.
A real peak around 15 years of age.
***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:
Use a log scale to see all the data when it varies over orders of magnitude.
***

### Transforming Data
Notes:
```
summary(pf$friend_count)
summary(log10(pf$friend_count + 1))
```
***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
#http://bit.ly/gridExtra
install.packages("gridExtra")
library(gridExtra)
p1 <- qplot(x=age, data=pf, 
            color=I('black'), fill=I('#F79420')) + 
  scale_x_continuous(breaks=seq(10,90,5), limit=c(10,90))
p1log <- qplot(x=age, data=pf, ylab="Log10 Count",
               color=I('black'), fill=I('#F79420')) + 
  scale_y_log10() + 
  scale_x_continuous(breaks=seq(10,90,5), limit=c(10,90))
p1sqrt <- qplot(x=age, data=pf, ylab="Sqrt Count",
                color=I('black'), fill=I('#F79420')) + 
  scale_y_sqrt() + 
  scale_x_continuous(breaks=seq(10,90,5), limit=c(10,90))
grid.arrange(p1, p1log, p1sqrt)
```

***


### Frequency Polygons

```{r Frequency Polygons}

```

***

### Likes on the Web
Notes:

```{r Likes on the Web}

```


***

### Box Plots
Notes:

```{r Box Plots}

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!