Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:
Observed that actual and perceived were not very corelated.
Also could see banding of perceived where people are guessing round numbers.
***

### Scatterplots
Notes:

```{r Scatterplots}
setwd('/home/jay/classes/R.class')
#install.packages('ggplot2')
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep='\t')
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point() +
  xlim(13, 90)
```

***

#### What are some things that you notice right away?
Response:
Younger people have way more friends. 
There is a spike at the older end of the spectrum as well.

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20) +
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:
Still a strong spike around age 20. It is perhaps close to normally distributed 
between 13 and 25.
There is still that "fake age" spike at 69.
***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha = 1/20) +
  xlim(13, 90)
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) +
  coord_trans(y="sqrt") +
  geom_jitter(alpha = 1/20) +
  xlim(13, 90) +
  ylim(0, 5000)
```

#### What do you notice?
Two modes, one around 25 and one around 55.
***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friend_count), data = pf) +
  coord_trans(y="sqrt") +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13, 90)
```

```{r friendships_initiated and age}
ggplot(aes(x = age, y = friendships_initiated), data = pf) +
  coord_trans(y="sqrt") +
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13, 90)
```
***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
#install.packages('dplyr')
library(dplyr)
age_groups <- group_by(pf, age)
# new data frame
pf.fc_by_age <- summarise(age_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) +
  #coord_trans(y="sqrt") +
  geom_line() +
  xlim(13, 90)
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) +
  xlim(13, 90) +
  geom_point(alpha = 1/20, position = position_jitter(h = 0), color = 'orange') +
  coord_trans(y="sqrt") +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, probs=.1, linetype=2, color='blue') +
  geom_line(stat = 'summary', fun.y = quantile, probs=.9, linetype=2, color='red') +
  geom_line(stat = 'summary', fun.y = quantile, probs=.5, color='blue')
```

#### What are some of your observations of the plot?
Response:
Median and mean are relatively close together.
***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count, alternative="less", method="pearson")
```
Got a correlation of -0.027. Mo meaningful correlation.
A correlation > 0.3 or < 0.3 is meaningful.

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:
likes_received (y) vs. www_likes_received (x)
```{r}
summary(pf)
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point(alpha = 1/5, color = 'orange') +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) +
  geom_smooth(method = "lm", color="red")
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(pf, cor.test(www_likes_received, likes_received))
```

Response:

***

### Moira on Correlation
Notes:
Regression assumes independent variables. If any are highly corelated  it can make 
it difficult to tell what is really driving the outcome.
***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell
names(Mitchell)
head(Mitchell)
summary(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x=Month, y=Temp), data=Mitchell) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
0

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
with(Mitchell, cor.test(Month, Temp))
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x=Month, y=Temp), data=Mitchell) +
  geom_point() +
  scale_x_discrete(breaks=seq(0,max(Mitchell$Month),12))
```

***

### A New Perspective

What do you notice?
Response:
A cyclical pattern.

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
summary(pf)
str(pf)
pf$age_with_months <- pf$age + (1 - (pf$dob_month / 12))
```

***

### Age with Months Means

```{r Age with Months Means}
age_with_months_groups <- group_by(pf, age_with_months)
# new data frame
pf.fc_by_age_months <- summarise(age_with_months_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())
pf.fc_by_age_months <- arrange(pf.fc_by_age_months, age_with_months)
head(pf.fc_by_age_months)
```

Programming Assignment
```{r Programming Assignment}
summary(pf.fc_by_age_months)
ggplot(aes(x = age_with_months, y = friend_count_mean), data = pf.fc_by_age_months) +
  geom_line() +
  xlim(13, 71) +
  ylim(0, quantile(pf.fc_by_age_months$friend_count_mean, 0.95))
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
p1 <- ggplot(aes(x = age, y = friend_count_mean), 
             data = subset(pf.fc_by_age, age < 71)) +
        geom_line() +
        geom_smooth()
p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean), 
             data = subset(pf.fc_by_age_months, age_with_months < 71)) +
        geom_line() +
        geom_smooth()
p3 <- ggplot(aes(x = round(age / 5) * 5, y = friend_count), 
             data = subset(pf, age < 71)) +
        geom_line(stat = "summary", fun.y = mean)
library(gridExtra)
grid.arrange(p2, p1, p3, ncol = 1)
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:
Scatter shows relation between two.
Look at different scales and add summary statistics.
Are ther correlated? What are the limits of looking at correlation?
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

