Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***
group_by example
```{r}
head(mtcars)
by_vs_am <- group_by(mtcars, vs, am)
head(by_vs_am)
by_vs <- summarise(by_vs_am, n = n())
groups(by_vs)
summarise(by_vs, n = sum(n))
```

```{r}
setwd('/home/jay/classes/R.class')
pf <- read.delim('pseudo_facebook.tsv')
head(pf)
suppressMessages(library(dplyr))
data <- subset(pf, select=c(age, gender, friend_count))
data <- subset(data, !is.na(gender))
by_age_by_gender <- group_by(data, age, gender)
pf.fc_by_age_gender <- summarize(by_age_by_gender, 
                mean_friend_count=mean(friend_count),
                median_friend_count=median(as.numeric(friend_count)),
                n=n())
head(pf.fc_by_age_gender)

ggplot(aes(x = age, y = median_friend_count), data = pf.fc_by_age_gender) +
  #coord_trans(y="sqrt") +
  geom_line(aes(color=gender))# +
  xlim(13, 90)
```

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_histogram()
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}

```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')
head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
library(ggplot2)
ggplot(aes(x=age, y=female / male), data=pf.fc_by_age_gender.wide) +
  geom_line() +
  geom_hline(linetype=2, yintercept=1)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}

```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined <- floor(2014 - pf$tenure / 365)
summary(pf$year_joined)
table(pf$year_joined)
pf$year_joined.bucket <- cut(pf$year_joined, breaks=c(2004, 2009, 2011, 2012, 2014))
table(pf$year_joined.bucket)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x=age, y=friend_count),
           data=subset(pf, !is.na(pf$year_joined.bucket))) +
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=median)

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x=age, y=friend_count),
           data=subset(pf, !is.na(pf$year_joined.bucket))) +
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=mean) +
  geom_line(linetype=2, stat = 'summary', fun.y = mean)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
pf.tenure <- subset(pf, pf$tenure >= 1)
pf.tenure$friend_rate <- pf.tenure$friend_count / pf.tenure$tenure
summary(pf.tenure)
```

***

### Friendships Initiated
Notes:

What is the median friend rate?
.2205

What is the maximum friend rate?
417

```{r Friendships Initiated}
ggplot(aes(x=tenure, y=friendships_initiated / tenure),
           data=subset(pf, tenure >= 1)) +
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

Using a smoother instead of binning.
```{r}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket))
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

Get rid of noise by binning x-axis.

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

